ASTR480 Final Report:\\ \textit{Asteroids in TESS}

### 2024-10-02 FinalReport/FinalReportDraft.tex
added figures to the report

A comparison between the two light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
\autoref{Fig:DifFlux} shows these two light curves for a chosen asteroid, ``Ruff''.
This was picked as it has a high number of points matched to detections, but of course still more interpolated points total.
There seems to be a systematic offset between the fluxes, with the interpolated points having consistently lower flux, (here the means differ by 58 counts)  %!check

@@

@@

A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
TESS should have rather evenly spaced data, as the FFIs are produced on a set candence, but sometimes data will be missing.
For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast Fourier transform or a direct least squares fit.

@@

@@

\section{Results}\label{Sec:Res}
\section{Discussion}\label{Sec:Disc}
\section{Conclusion}\label{Sec:Conc}

\section{Figures}
%*Figures: Only here to keep them out of the way. Will be placed in text as it grows

\begin{figure}
  \centering
  \includegraphics[width =\columnwidth]{../Test Code/Testing Figures/differentFluxes Ruff .pdf}
  \caption[Light curves of Ruff]{The light curves of the asteroid ``Ruff'', with the flux of the interpolated positions in blue circles and the flux from the matches to detected sources in orange diamonds.
    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in Modified Julian Date.}
  \label{Fig:DifFlux}
\end{figure}

>>>>

@@

\begin{figure}
  \centering
  \includegraphics[width =\textwidth]{../OzData/SingleBodyLC1991 VT.pdf}
  \caption[Light curves of 1991 VT]{The light curves of the asteroid 1991 VT throughout its entrie journey across sector 29, with the flux of the interpolated positions in blue circles and the flux from the COM in orange diamonds, and the \texttt{TESSELLATE} flux for any matched points in the black squares.
    The Flux axis is measured in counts, as calculated by \texttt{TESSreduce}, and the times are in modified Julian date (MJD).}
  \label{Fig:FullSingleLC}
\end{figure}
A comparison between the three light curves is interesting, as one might expect only the brightest parts of the lightcurve will be matched.
\autoref{Fig:FullSingleLC} shows these three light curves for a single body.
The COM flux seems much more stable than the interperted flux on its own, and at a higher average flux count.
This should result in more accurate periodograms as there is less overall curve to the data.
The matched points are comparatavely sparce for this object, which is not the case for all.
Many have no matches, and some have almost fully matched, this will be discussed more later.

@@

A widely used technique in determining the periods of astronomical data is the Lomb-Scargle periodogram \citep[\citet{Lomb1976,Scargle1982}, but see][for a review]{VanderPlas2018}.
This method can efficiently take discrete timeseries data of which the observations are unevenly spaced and calculate the frequencies present in the sample.
TESS should have rather evenly spaced data, as the FFIs are produced on a set cadence, but sometimes data will be missing.
For example, it may have been sigma clipped due to a spike in brightness that is unphysical for an asteroid, or TESS sees the same asteroid in the same sector on either side of the day long mid-sector break it takes for data transfer.
The Lomb-Scargle periodogram deals with this uneven sampling much better than either a fast Fourier transform or a direct least squares fit.

@@

%TODO single object periodogram
%Maybe a bad one too

@@

\section{Results}\label{Sec:Res}

This is the results for TESS sector 29.
This sector was chosen because it is early in the \qty{10}{\min} cadence of TESS FFIs, and had been fully run through TESSELLATE.
Similar results can be found in any of the sectors, just the computaion time was a limiting factor.

\subsection{Asteroid Properties}\label{SubSec:AstPropRes}

The orbital element properties of the asteroids were returned by JPL Horizons queries.
A plot of the semi-major axis against the other orbital elements, eccentricity and inclination, is shown in \autoref{Fig:AEI}.
The disticntion between the resonances in the main belt are clearly seen. There are lines seperating the middle belt from the inner and outer belt.
The Jupiter trojans are well defined, at just past \qty{5}{\au}, with a small range of eccentricity but a large spread in the inclination.
Inbetween the outerbelt and the Jupiter trojans, sit some Mars crossers.
A some near earth objects (NEOs) at smaller $a$ and a few more objects that are scattered with high $e$ and or $i$ complete the distrabution.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../OzData/AEIplot29.pdf}
  \caption[aei distribution]{a, e, i, plot of all the objects on sector 29}
  \label{Fig:AEI}
\end{figure}

The number the belong to each class of objects is highlighted in \autoref{Fig:NumPerClass}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../OzData/classesBar29.pdf}
  \caption[Class distribution]{A distrabution of the number of asteroids per class that \texttt{SkyBot} knows the object is. }
  \label{Fig:NumPerClass}
\end{figure}

\subsection{Detectability}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../OzData/recoverdHistBkgLimof10.pdf}
  \caption[Recovery plot]{A histogram of known average visual magnitude of each object against the whether or not the object has a lightcurve with an average flux value above 10 %!CHECK
    counts for its COM flux.
  }
  \label{Fig:RecovOverBkg}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../OzData/pointsMatchesNumberHistTEST.pdf}
  \caption[short]{A histogram of the number of point in the lightcurve of each asteroid (blue) against the number of matches each asteroid gets in the \texttt{TESSELLATE} detections(orange dashed). }
  \label{Fig:MatchInterpHists}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{../OzData/matchesHist29.pdf}
  \caption[Magnitude of matches]{The visual magnitude of the asteroid against the fraction of asteroids in this bin with more than 3 %!CHECK
    matches.
  }
  \label{Fig:MagofMatches}
\end{figure}

\section{Discussion}\label{Sec:Disc}
\section{Conclusion}\label{Sec:Conc}
